\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\usepackage{tikz-cd}
\renewcommand{\d}{\mathrm{d}}

\begin{document}

\title{Math 110AH Homework 8}
\author{Nathan Solomon}
\maketitle

\textbf{Assignment due November 29th at 11:59 pm.}

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Let $G$ act on a set $X$. Prove that if $x,y \in X$ satisfy $ax=y$ for some $a \in G$ then $G_y = a \cdot G_x \cdot a^{-1}$.
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Let $G$ be a group, $a \in G$. Show that the number of elements in the conjugacy class $\{ bab^{-1}, b \in G \}$ divides $|G|$.
}}}\bigskip

Let $H$ be the conjugacy class of $a$. Since $H$ is normal in $G$, the quotient $G/H$ is a group. By Lagrange's theorem,
FIND A SOLUTION THAT WORKS FOR INFINITE GROUPS TOO

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            An action of a group $G$ on a set $X$ is called \textit{double transitive} if for any two pairs $(x_1, x_2)$ and $(y_1, y_2)$ of elements of $X$ such that $x_1 \neq x_2$ and $y_1 \neq y_2$ there is $a \in G$ such that $ax_1 = y_1$ and $ax_2 = y_2$. Prove that $|G| \geq |X|^2 - |X|$. (Hint: Consider an action of $G$ on $X \times X$.)
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            \begin{itemize}
                \item (a) Let $H$ be a subgroup of a finite group $G$. Prove that the number of different conjugate subgroups $xHx^{-1}, x \in G$, is at most $[G:H]$.
                \item (b) Let $H$ be a subgroup of a finite group $G$. Prove that if $G$ is the union of $xHx^{-1}$ over all $x \in G$, then $H = G$.
                \item (c) Let a finite group $G$ act transitively on a set $X$ consisting of at least two elements. Prove that there exists a $g \in G$ fixing no element of $X$.
            \end{itemize}
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Determine all Sylow $p$-subgroups of $A_5$.
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Find the number of all Sylow $p$-subgroups of $S_p$ ($p$ is prime).
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Prove the following Useful Counting Result. Let $H$ be a \textit{proper} subgroup of a finite group $G$. Suppose that $|G|$ does not divide $[G:H]!$. Then $G$ contains a nontrivial normal subgroup $N$ such that $N$ is a subgroup of $H$. In particular, $G$ is not simple.
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Prove that all groups of order $2p^n$ ($p$ is prime, $n \geq 1$) are not simple.
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            \begin{itemize}
                \item (a) Let $H \subset G$ be a subgroup. Prove that if $H$ is contained in the center of $G$ and the quotient group $G/H$ is cyclic, then $G$ is abelian.
                \item (b) Prove that any group of order $p^2$ is abelian ($p$ is prime).
            \end{itemize}
}}}\bigskip

\section{}
\noindent\fbox{\fbox{\parbox{6.5in}{
            Prove that $\{e\}$, $S_n$, and $A_n$ are the only normal subgroups in $S_n$ if $n \geq 5$.
}}}\bigskip

\end{document}
