\documentclass[class=article, crop=false]{standalone}
\input{../preamble.tex}

\fancyhf{}
\lhead{Nathan Solomon}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
\section{4/18/2024 lecture}
THE MIDTERM WILL BE MAY 9TH IN CLASS (12:30 TO 1:45). IT WILL CONSIST OF 50 MULTIPLE CHOICE QUESTIONS.
\par
HOMEWORK FOR THIS CLASS WILL BE POSTED SOON.
\par
Meteorites are chunks of rock (from meteors) that land on Earth. Potassium-argon dating shows that most meteorites formed around 4.6 billion years ago.
\par
A photon with energy 1 eV is infrared, 3 eV is visible, and 5 eV is ultraviolet. A photon needs a minimum energy of 13.6 eV to ionize a hydrogen atom. X-rays have energies of 100 eV to 100 keV per photon.
\par
There are three terms we use to descripe spectra: emission, continuous, and absorption. The continuous spectra tell us temperature, whereas the emission and absorption spectra tell us chemical composition. We observe absorption spectra when a cloud blocks light, and emission spectra when a cloud absorbs and re-emits light.
\par
Hotter objects emit more energy AT EVERY FREQUENCY than colder objects (assuming they have the same absoptivity/emissivity). Hotter objects emit photons with a higher average energy (higher average frequency). The Stefan-Boltzmann law says luminosity per square meter is proportional to temperature to the fourth power. Wein's law says the peak wavelength is inversely proportional to the temperature. Also, the peak frequency is proportional to temperature, but be careful because the peak wavelength and peak frequency do not correspond to each other.
\par
Reflecting telescopes are much more powerful than classical telescopes.
\par
Launching telescopes into space is expensive, but then you don't have to worry about air turbulence or about scattering from the atmosphere (including from clouds and dust). Some fequencies of light, such as x-rays, cannot penetrate the atmosphere, so those telescopes need to be launched into space.
\par
Adaptive optics is a recent technology, in which mirrors within a telescope move hundreds of times a second to cancel out effects like air turbulence.
\par
Interferometry is a technique which combines two or more telescopes that are far apart to obtain a better angular resolution. For a traditional telescope, angular resolution (the smallest angular size you can see) is proportional to the wavelength of the light divided by the diameter of the telescope, but for telescope arrays using interferometry, the ``diameter" is the distance between the telescopes. This is especially useful for radio waves, since the large wavelength would mean the resolution is large (large is bad).

\end{document}
