\documentclass[class=article, crop=false]{standalone}
\input{../preamble.tex}

\fancyhf{}
\lhead{Nathan Solomon}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
\section{5/2/2024 lecture}
\subsection{Definition of ``life"}
Life has all of the following:
\begin{itemize}
    \item Order -- living cells and bodies are organized, not random
    \item Reproduction -- although mules are considered alive, and viruses aren't
    \item Growth \& development
    \item Energy utilization -- necessary to prevent their entropy from increasing
    \item Response to environment
    \item Evolucionary adaptations
\end{itemize}
\subsection{Chemical composition of life on Earth}
The Miller-Urey experiment zapped a bunch of water, methane hydrogen, and ammonia to create complex amino acids. This supports the theory that life could have spontaneously arisen from the primordial soup.
\par
60\% of the world's iron ore arises in banded formations, dating back to when the first cyanobacteria appeared. That's because there used to be a ton of iron dissolved in the oceans, then 2.5 million years ago, cyanobacteria in the oceans produced oxygen, which combined with the iron to form compounds like rust. The iron then sank to the ocean floor and sedimented. The oxygen level in the atmosphere took over a billion years to actually reach the level it's at now (roughly 20\%). Once there was enough oxygen for aerobic respiration to occur, eukaryotes popped off.
\par
The last common ancestor of all life on Earth is estimated to have existed around 4 billion years ago.
\par
All 36 modern phyla of animals, such as chordata and arthropoda, appeared between 542 and 500 million years ago, at the beginning of the Cambrian explosion.
\par
By mass, humans are
\begin{itemize}
    \item 65\% oxygen
    \item 18.5\% carbon
    \item 9.5\% hydrogen
    \item 3.3\% nitrogen
    \item 1.5\% calcium
    \item 1\% phosphorus
    \item .4\% potassium
    \item .3\% sulfur
    \item .2\% sodium
    \item .2\% chlorine
    \item .1\% magnesium
    \item Other trace elements, which may still be important
\end{itemize}
\subsection{Professor Rich's dogs}
Professor Rich has two dogs: Holly and Maggie. Maggie likes to poop in the house.
\par
This will be on the final.

\end{document}
