\documentclass[class=article, crop=false]{standalone}
\input{../preamble.tex}


\fancyhf{}
\lhead{Nathan Solomon}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
\section{5/15/2024 lecture}
Last lecture, we discussed how any open neighborhood of the identity in a connected Lie group generates the whole group.

\subsection{The exponential map}
The exponential function $\exp: \C \rightarrow \C$ is defined by
\[ \exp(x) = \sum_{n=0}^\infty \frac{x^n}{n!}. \]
Since it's defined as a power series, it is an \textit{entire} function.
\par
Alternatively, the exponential function on $\C$ can be defined as the unique solution to
\[ f'(x)=x, \hspace{1cm} f(0)=1. \]
For any $n \times n$ matrix over either $\R$ or $\C$, we can define $\exp$ with the same power series.
\begin{prop}
    For any $A \in \C^{n \times n}$, the following power series is absolutely convergent:
    \[ \exp(A) := \sum_{k=0}^\infty \frac{A^k}{k!}. \]
\end{prop}
\begin{proof}
    \begin{align*}
        \abs{\exp(A)_{i,j}} &= \abs{\left( \sum_{k=0}^\infty \frac{A^k}{k!} \right)_{i,j}} \\
                            &\leq \sum_{k=0}^\infty \frac{1}{k!} n^{k-1} \max_{p,q} \abs{A_{p,q}}^k
    \end{align*}
    So every entry of $\exp(A)$ is bounded by an absolutely convergent series.
\end{proof}
\begin{example}
    Let $A$ be any 2 by 2 complex matrix such that $A^2=I$. Then $\exp(tA)= \left( \cosh t \right) I + \left( \sinh t \right) A$. Notably, if $t=i \theta$, then $\exp(i \theta A) = I \cdot \cos \theta + A \cdot i \sin \theta$.
\end{example}
\subsection{The matrix logarithm}
The matrix logarithm of $A$ can be defined by the following formula:
\[ \log (I+A) = \sum_{k=1}^\infty \frac{A^k}{k} (-1)^{k+1}. \]
We say that the log of a matrix exists iff that series is absolutely convergent -- for example, if the largest eigenvalue of $A$ has magnitude less than 1, then the matrix log exists.
\begin{thm}
    \begin{enumerate}
    \item If the log of a (real or complex) matrix exists (meaning $\log A$ is an absolutely convergent power series), then
    \begin{align*}
        \log(\exp(A))&=A \text{    and} \\
        \exp(\log(A))&=A.
    \end{align*}
\item $\exp(0)=I$ and $\log(I)=0$.
\item The following statements are equivalent:
    \begin{itemize}
        \item $AB=BA$
        \item $\exp(A+B)=\exp(A)\exp(B)$
        \item $\log(AB)=\log(A)+\log(B)$
    \end{itemize}
    DOES THAT STILL WORK EVEN IF THE LOG DOESNT CONVERGE??? ARE THOSE THREE STATEMENTS REALLY ALL EQUIVALENT???
    \end{enumerate}
\end{thm}
\begin{proof}
    The proofs of these are an exercise.
    \par
    \[ \exp(A+B)=\sum_{k=0}^\infty \frac{1}{k!} \sum_{\ell=o}^\infty \binom{k}{\ell} A^\ell B^{k-\ell} \]
    compare that to $\exp(A)\exp(B)$.
\end{proof}
\begin{thm}
    The exponential of any $A \in \C^{n \times n}$ is invertible, and its inverse is the exponential of $-A$.
\end{thm}
\begin{proof}
    \[ \exp(A)\exp(-A)=\exp(0)=I \]
\end{proof}
\begin{prob}
    Is the matrix exponential from the set of complex $n \times n$ matrices a surjection onto the set of invertible complex $n \times n$ matrices?
\end{prob}
\begin{prop}
    For any complex $n \times n$ matrices $A$ and $B$, there exists a matrix $C$ such that
    \[ \exp(A) \exp(B) = \exp(C). \]
\end{prop}
\begin{proof}
    For some very small $t \in \C$,
    \[ \exp(tA)\exp(tB)-\exp(t(A+B))= \frac{t^2}{2} [A,B]+O(t^3). \]
\end{proof}
INCLUDE THE BAKER CAMPBELL HAUSDORFF FORMULA. It gives a formal power series in $t$ (not necessarily convergent) for $\log(\exp(tA)\exp(tB))$ in terms of iterated commutators of $A$ and $B$
SEE MCGREEVYS NOTES ABOUT HOW TO DERIVE THAT USING A DIFFERENTIAL EQUATION AND AN ANSATZ
\subsection{Suzuki-Trotter formula}
\begin{thm}
    \[ e^{A+B}=\lim_{n \rightarrow \infty} \left( e^{A/n}e^{B/n} \right)^n. \]
\end{thm}
\begin{proof}
    \begin{align*}
        e^{tA/n}e^{tB/n}&=\exp \left( \frac{t}{n} (A+B)+ \frac{t^2}{2n^2} [A,B] + O\left( \frac{t^3}{n^3}\right) \right)
    \end{align*}
    FINISH THIS PROOF
\end{proof}
\subsection{Lie algebras}
If $G$ is a topologically closed subgroup of $GL(n,\F)$ (where $\F$ is either $\R$ or $\C$), let $\mathfrak{G}= \left\{ A \in \F^{n \times n}: e^{tA} \in G \forall t \right\}$. Then
\begin{enumerate}
    \item $\mathfrak{G}$ is a vector space
    \item $\mathfrak{G}$ is closed under the Lie bracket (which for matrix Lie algebras is always the commutator, I think??? IS THAT TRUE?)
    \item $\mathfrak{G}$ is the tangent space to $G$ at the identity
    \item $G$ is a Lie group
\end{enumerate}
To prove statements (1), you can use the Suzuki-Trotter formula. THE REST OF THESE STATEMENTS ARE EXERCISES

\end{document}
