\documentclass[class=article, crop=false]{standalone}
\input{../preamble.tex}


\fancyhf{}
\lhead{Nathan Solomon}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
\section{4/2/2024 lecture}

\subsection{Syllabus review}

The grading scheme is as follows:
\begin{itemize}
    \item Weekly discussion sections are 20\% of final grade (2\% per day)
    \item Weekly online writings are worth 10\% of your grade
    \item Exams are each worth 15\% of your grade
    \item The papers are worth 40\% of your grade
\end{itemize}
Exams are open book and open note take-home exams. The midterm is May 3rd, and the Final is June 13th.
\par
The ``paper" portion of your grade is base on a ``lyrics anaysis" due April 26th, which is 10\% of your grade, a ``musical analysis" due May 17th, which is also 10\%, and a 5-6 page final paper due May 31st, which is 20\% of your grade.
\par
This classes is not graded on a curve.
\par
The recommended (but not required) book is Jonathan Gould's \emph{Can't Buy Me Love}.

\subsection{Mop Tops}
The Beatles all wore a hairstyle that was very different from the norm. At the time, all men wore very short hair. When they were young, they had hair that looks fairly short to us, but looked overgrown to everyone at the time, hence the name ``mop tops".

\subsection{The Quarrymen}
From 1956-1959, John Lennon played as part of a band called The Quarrymen. They learned how to play by listening to American records.

\subsection{History}
From 1960-1962, The Beatles played as a bar band in Hamburg \& Liverpool. They were extremely popular amongst teenagers, but the media didn't take them seriously at first. In 1961, they changed their name from The Quarrymen to The Beatles.
\par
Every single that they released from 1963-1966 was a top hit. During that time, they worked extremely hard recording, touring, and making films.
\par
In 1964, they played in America for the first time, on \emph{The Ed Sullivan Show}.
\par
From 1966 to 1970, they stopped touring, and did \textbf{studio recording only}. This allowed them to make much more experimental music.
\par
All of these phases of their career were a result of external circumstances. Therefore when analyzing The Beatles, we need to consider the constraints on their play style, and how their music evolved as a result of those constraints.

\end{document}
